\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2020-12-14 popl-2019-journal-v/proving-full-abstraction.tex
Add file with JFP reviews in order to collect responses, fix trivial typos already.

\begin{jversion}
Assumptions about the system that relatedness is predicated on are gathered in (Kripke) worlds.
To a first approximation, a world is a semantic model of the memory.
In its simplest form, it is a collection of invariants that the memory must satisfy.
The invariants of a world can vary in complexity and expressiveness depending on the application.
The possible contents of the memory also influences what the world looks like.
For instance, the uniqueness of the linear capabilities on \trgcm{} and \srccm{} is modelled by the worlds using a form of memory ownership assumptions.
In order to relate \trgcm{} and \srccm{}, we model all the features of \srccm{} in the world which means we have to model:
\begin{itemize}
\item Three kinds of memory: heap, stack of local frames, and free stack
\item Linearity
\item Call stack
\item Seals
\end{itemize}
The three kinds of memory are modelled by having three sub-worlds where each sub-world is its own little world in a traditional sense.
Linearity is modelled by adding ownership to certain parts of the world that capabilities can take ownership of ensuring that they are the sole reference for that part of memory.
Memory satisfaction (the relation that decides whether two memories are related in the world) models the call stack by ensuring that the memory is actually shaped like a stack.
Finally, the seals are modelled by seal invariants that make sure that seals are only used on permitted sealables.
In the following, we present the world and go into details about how each of the four features are modelled.

@@

%%% spatial/spatial_owned
We cannot let multiple linear capabilities reference the same memory.
The spatial region represent ownership of a part of memory which ensures that a linear capability is not aliased.
Even though spatial regions gives the right to reference part of memory, we still need the world to specify the remainder of the memory that may be referenced by linear capabilities.
To this end, we have shadow regions.
A shadow region is a shadow copy of a spatial region in the sense that it specifies part of memory, but it does not give the right to reference that part of memory.
This does not mean that the memory is necessarily not referenced.
A compatible world may claim a shadow region as spatial which allows the other world to reference the memory (we return to the notion of compatible worlds w.r.t. ownership in Section~\ref{subsubsec:joining-worlds}).
Specifically, we add tags $\spatialo$ and $\spatial$ to the spatial regions:
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \cup \\ 
    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg})  \\
  \end{array} \right.
\]
For readability, we also add a tag $\pure$ to the shared regions:
\[
  \Regionh = \{\pure \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) 
\]

% What parts of memory are linear
We will extend the regions further in Sections~\ref{subsubsec:seals}~and~\ref{subsubsec:ft-and-revocation}, but for now we continue the definitions of the three sub-worlds.
The sub-world $\Worldh$ specifies the heap memory which can be referenced by both linear and normal capabilities, so it should contain both shared and spatial regions.
For this reason, it is defined as
\[
  \Worldh = \RegName \parfun (\Regionh \cup \Regions)
\]
\srccm{} internalizes the \stktokens{} stack, so it can only be referenced by linear capabilities which means that the two stack regions should only have spatial regions.
For instance the $\Worldfs$ is defined as
\[
  \Worldfs = \RegName \parfun \Regions
\]
$\Worlds$ can also only be referenced by linear capabilities, so it should also only have spatial regions.
$\Worlds$ not only models the memory contents of the local stack frames, it also models the call stack that consists of the stack frames.
Conceptually, this means that each of the stack frames is connected with a return point in some code.
In a traditional C calling convention, the code return point would even be stored in the stack frame.
However, with \stktokens{} a stack frame does not contain any information about the corresponding code return point.
Instead, the stack frame and code return point is connected by the capabilities that reference them as they are sealed with the same seal and together then constitute the sealed return pair.
In order to ensure that each call actually returns to the correct point of the code, we must still include the address of the return point in our model.
To this end, each shared region in $\Worlds$ must be paired with a return address:
\[
\Worlds = \RegName \parfun (\Regions \times \Addr)
\]

% Ensuring linearity
The spatial regions add the necessary bookkeeping to the worlds to model linear capabilities.
The logical relation presented in Section~\ref{subsec:logical-relation} uses this bookkeeping to ensure that linear capabilities \emph{uniquely} references part of memory.

Given a world, we want to be able to express that a capability, that is otherwise valid with respect to the world, is not linear or indirectly depends on a linear capability.
This is expressed by stripping the world of all its ownership which corresponds to replacing all $\spatialo$ regions with $\spatial$ regions and then requiring that the capability is valid w.r.t.\ this new world.
We refer to this as the shared part of the world and define the function $\purePart{}$ which turns all spatial regions into shadow copies.
\begin{definition}[The shared part of a world]
  \label{def:purePart}
  For any world $W$, we define
  \begin{align*}
    \purePart{W} &\defeq (\purePart{\pwheap},\purePart{\pwpriv},\purePart{\pwfree})\\
    \purePart{W_\var{heap}} &\defeq \lambda r\ldotp
                       \begin{cases}
                         (\spatial,\var{H}) & \text{if } W_\var{heap}(r) = (\spatialo,\var{H})\\
                         W_{\var{heap}}(r) & \text{otherwise}
                       \end{cases}\\
    \purePart{W_\var{call\_stk}} &\defeq \lambda r\ldotp
                       \begin{cases}
                         ((\spatial,\var{H}),\opc) & \text{if } W_\var{call\_stk}(r) = ((\spatialo,\var{H}),\opc)\\
                         W_\var{call\_stk}(r) & \text{otherwise}
                       \end{cases}\\
    \purePart{W_\var{free}} &\defeq \lambda r\ldotp
                       \begin{cases}
                         (\spatial,\var{Hs}) & \text{if } W_{\var{free}}(r) = (\spatialo,\var{Hs})\\
                         W_{\var{free}}(r) & \text{otherwise}
                       \end{cases}
  \end{align*}
\end{definition}
\subsubsection{Seals}
\label{subsubsec:seals}
So far, the world represents a collection of assumptions on the memory contents that value correctness may depend on.
However, value correctness may depend on other assumptions.
Specifically, \stktokens{} has certain assumption on the seals used for return capabilities and closures.
For instance, a return seal must only be used to seal the return pointer of one specific return point.
Therefore, in addition to a relation on memory segments, some regions also carry a \emph{seal interpretation function} that relates the sealables that may be sealed with a given seal.
  \[
    \Seal \parfun \World \fun \Rel{\SealableCaps \times \SealableCaps}
  \]
%% Shared region has a seal indexed invariant that specifies what sealables can be sealed with a specific seal
  In \stktokens{}, once a seal has been used for a specific purpose (e.g.\ for sealing return capability pairs for a specific call site), it can never be reused for a different purpose.
  This is because there may still be copies of return capabilities out there, signed with the seal.
  This situation is similar to the situation for non-linear memory capabilities, so we only allow shared regions to carry seal interpretation functions, as we will see that those regions can never be revoked in future worlds.
\begin{multline*}
  \Regionh = 
  \{\pure \} \times (\World \fun \Rel{\MemSeg\times\MemSeg}) \times \\
  (\Seal \parfun \World \fun \Rel{\SealableCaps \times \SealableCaps})
\end{multline*}
We also refer to the seal interpretation function as the seal invariant, and we will refer to the memory relation as the memory invariant or just invariant when it is unambiguous.

@@

Safety of capabilities should be defined, so it is monotone with respect to the future world relation.
This means that the system assumptions that capabilities rely on for safety are defined such that capabilities remain safe during execution.
The same holds for the memory invariants stored in the world: capabilities stored in memory should remain valid under allowed changes in system assumptions.
This means we require the world-indexed memory invariants to be monotone in the world.
In other words, a pair of memory segments that are related now will stay related in future worlds.
This changes the spatial regions as follows:
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \cup \\
    \{\spatialo \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \cup \\
    \{\revoked \}
  \end{array} \right.
\]
We make a similar change to $\Regionh$. The seal invariants are monotone as well.

@@

The stack relation also partitions the authority of the world, so it can relate the stack frames in a way that respect linearity.

@@

So far, we have talked about the logical relation as though we define one.
However, we actually define two logical approximations that only differ in the observation relation.
We define a \srccm{} configuration to logically approximate a \trgcm{} configuration when the halting termination of the \srccm{} configuration implies the halting termination of the \trgcm{} configuration.
This also means that \srccm{} configurations that terminates by failing termination are related to any \trgcm{} configuration.
Intuitively, this is because the $\failed$ configuration signals that there was an attempt to break the guarantees of the capability machine.
For instance, a piece of code could have attempted to read from a part of memory it does not have access to, or a callee could have attempted to return out of order.
In both cases, we haven't defined a way to recover from such attempts to break the guarantees, so we are content with failure.
\begin{multline*}
  \lro[\preceq,(\ta,\stkb,\_,\_)] \defeq \\\left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
    \begin{array}{l}
      \forall i \leq n \ldotp \\
      \quad \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta,\stkb} \\\qquad\Rightarrow (\ms_T,\reg_T) \term\\
    \end{array}
\right\}
\end{multline*}
The step-indexing plays a role here because we are only interested in \srccm{} configurations that terminate in $n$ or fewer steps.
However, if the \srccm{} configuration terminates successfully in $n$ steps, then the \trgcm{} configuration should just terminate in any number of step (possibly more than $n$ steps).
For the most part, it would make sense to require the \trgcm{} configuration to terminate in the same amount of steps as the \srccm{} configuration as they run in lockstep for most of the computation.
However, when it comes to calls and returns, the two configurations stop running in lockstep.
The \srccm{} configuration handles calls and returns in one step whereas \trgcm{} configurations need to execute each instruction of the call preparation as well as the return code.

@@

In Figure~\ref{fig:value-relation}, we have sketched the two value relations.
This shows that for the most part, words on \srccm{} are related to words on \trgcm{} that are syntactical identical.
The only exception is stack pointers on \srccm{} that are related to linear capabilities on \trgcm{}.
Note that the return pointers of \srccm{} are not related to anything as it is never safe for any program, trusted or not, to have them.
The \srccm{} return pointers should only occur under a return seal, and they should only be used in a jump in which case the \srccm{} semantics transforms them to the capabilities they correspond to.

@@

In our result, we assume well-formed components which puts certain syntactic constraints on the components.
We also have the semantic assumption that trusted components are reasonable.
Both assumptions need to be captured in the logical relation in order for us to rely on them.
To this end, we define a \emph{code region} which captures the syntactic and semantic assumptions we make on components.
The memory invariant of the code region is defined as
\begin{multline*}
  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} =\\
  \left\{\npair{\arraycolsep=0pt\left(\array{l}\code \uplus \mspad,\\ \code \uplus \mspad\endarray\right)} \middle|
    \begin{array}{l}
    \dom(\code) = [\baddr,\eaddr] \wedge ([\baddr - 1, \eaddr + 1] \subseteq \ta
      \wedge\\
      \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss \wedge \trust = \trusted) \vee \\
      \quad ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
      \mspad = [\baddr-1 \mapsto 0] \uplus [\eaddr + 1 \mapsto 0]\wedge\\
      \sigrets,\sigcloss,\ta \vdash_{\mathrm{comp-code}} \code \wedge\\
      \forall a \in \dom(\code)\ldotp\\
      \quad\npair{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
    \end{array}
  \right\}
\end{multline*}
The code region is more restrictive than the standard region.
It only allows one memory segment, namely $\code$ padded with zeroes that make sure that two capabilities cannot be spliced to cause unintended control-flow.
We use the relation to reason about trusted components (well-formed and reasonable) as well as untrusted components (well-formed).
The assumptions we can make on the code depends on whether it is part of a trusted or untrusted component.
This is captured by requiring the contents of the code memory to be in the trusted or untrusted value relation depending on the trustworthiness of the code.
That is, if all the code memory addresses are in the trusted address space and the seals are from the global seals, then the component is trusted.
On the other hand, if the code memory addresses are disjoint from the trusted addresses and there are no return seals, then the component is untrusted.
In either case, the words should be in the value relation with respect to the $\purePart{}$ of the world which means that the code memory cannot contain linear capabilities.

@@

% Closure seals
Unlike return seals, both trusted and untrusted components can have closure seals.
For untrusted components (components with their code address space disjoint from the trusted address space), we allow everything in the untrusted value relation to be sealed.
Intuitively, untrusted components are assumed to have access to words from the untrusted value relation, and we cannot know how the words are used, so we need to assume that an untrusted component may seal untrusted words.
Trusted components only use closure seals for sealed capability pairs that represent actual closures.
The code capability for a closure must point to the code memory because it is the only part of memory that is executable.
Untrusted components cannot safely have a capability for a trusted components code (it could be used to read return capabilities or start execution in the middle of a call), so capabilities for the code memory of a trusted component is in the trusted value relation.
While it is not safe to give a bare capability for a trusted components code memory, it can be perfectly safe to give a sealed capability for a trusted components code.
For this reason, the seal invariant allows executable capabilities from the trusted value relation to be sealed with a closure seal.

@@

The world can be seen as an authority specification which means that it dictates what kind of capabilities can address a certain part of memory.
Specifically, linear capabilities can only address memory governed by a $\spatialo$ region, and normal capabilities can only address memory governed by a $\pure$ region.
All the permission based condition we define project the regions that the capability may address from the world.
The addressable $\addressable{}$ function takes care of the projection:
\[
  \addressable{\lin,W} \defeq
  \begin{cases}
    \{ r \mid W(r) = (\pure,\_) \} & \text{if $\lin = \normal$} \\
    \{ r \mid W(r) = (\spatialo,\_) \}  & \text{otherwise (i.e. $\lin = \linear$)} \\
  \end{cases}
\]

We capture the essence of what it means for a capability with read permission to be safe in the condition $\readCond[]{}$.
% each of the regions upper bounded by std region
The main purpose of $\readCond[]{}$ is to make sure that only safe words can be read from the memory governed by a read capability.
This is done by putting an upper bound on what requirements an invariant can impose on the memory segments governed by the capability.
In particular a region that governs the memory a read capability has access to can at most allow safe values to be read.
Without this requirement, a read capability could potentially be used to break memory invariants if it were used to read capabilities that has the authority to break memory invariants.
The read condition is defined as follows
\[
  \readCond{\lin,W} = \left\{ \npair{A} \middle| 
    \begin{array}{l}
      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
      \quad \exists R : S \fun \powerset{\nats} \ldotp\\
      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
      \qquad (\lin = \linear \Rightarrow \forall r \in S \ldotp |R(r)|  = 1) \wedge\\
      \qquad \forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H
    \end{array}
  \right\}
\]
% splicing and what part of the range each region governs
The $\readCond[]{}$ is compatible with all the operations that can be performed on capabilities.
This means that if two capabilities, for which $\readCond[]{}$ holds, are spliced together, we can establish that $\readCond[]{}$ holds for the resulting capability.
To support this, we require the presence of a set of regions $S$ that governs the addresses the capability has authority over rather than just a single region.
If we need to establish the $\readCond[]{}$ after a splice, we can simply use the union of the regions that witnessed the $\readCond[]{}$ of the two individual capabilities.
We also need to support splitting which is no problem for normal capabilities as the same $\pure$ region can be used to establish the $\readCond[]{}$ for multiple normal capabilities.
On the other hand, a $\spatialo$ region can only be used to establish the $\readCond[]{}$ for one linear capability because the ownership of a spatial region can only go to one world when splitting the ownership.
None the less, we need to support arbitrary splitting of linear capabilities, which means that $\readCond[]{}$ must make sure that the necessary regions are in the world to argue that the result of a split preserves $\readCond[]{}$.
This is why, $\readCond[]{}$ requires all regions to only govern one address when the capability is linear.
This means that after a split, the authority of the regions for the bottom half of the split can go to one capability and the remaining regions can go to the top half.

@@

Similarly to $\readCond[]{}$, we define a condition that captures the essence of what it means for a capability with writer permission to be safe.
We call this condition $\writeCond[]{}$.
A capability with write permission can be used to write to memory.
The question is, what can we safely allow to be written to memory without any memory invariants being broken.
The answer to this is anything - even words that are unsafe.
Say, you manage to write something that can break memory invariants, then it would not be possible to read it back again as write permission, generally speaking, does not entail read permission.
If the capability had read permission, then $\readCond[]{}$ would make sure that the word would have to be safe\footnote{It should not be possible to obtain a capability that can be used to break invariants. After all, if such a capability was obtained, memory invariants could be broken.
  However, the $\writeCond[]{}$ tries to capture the essence of safety and in principle it is safe to write an unsafe capability that cannot be read back.}.
It should always be possible to write safe values, so we impose this as a lower bound.

>>>>

\begin{jversion}
Assumptions about the system that relatedness is predicated on are gathered in (Kripke) worlds.
To a first approximation, a world is a semantic model of the memory.
In its simplest form, it is a collection of invariants that the memory must satisfy.
The invariants of a world can vary in complexity and expressiveness depending on the application.
The possible contents of the memory also influence what the world looks like.
For instance, the uniqueness of the linear capabilities on \trgcm{} and \srccm{} is modelled by the worlds using a form of memory ownership assumptions.
In order to relate \trgcm{} and \srccm{}, we model all the features of \srccm{} in the world which means we have to model:
\begin{itemize}
\item Three kinds of memory: heap, stack of local frames, and free stack
\item Linearity
\item Call stack
\item Seals
\end{itemize}
The three kinds of memory are modelled by having three sub-worlds where each sub-world is its own little world in a traditional sense.
Linearity is modelled by adding ownership to certain parts of the world that capabilities can take ownership of ensuring that they are the sole reference for that part of memory.
Memory satisfaction (the relation that decides whether two memories are related in the world) models the call stack by ensuring that the memory is actually shaped like a stack.
Finally, the seals are modelled by seal invariants that make sure that seals are only used on permitted sealables.
In the following, we present the world and go into details about how each of the four features are modelled.

@@

%%% spatial/spatial_owned
We cannot let multiple linear capabilities reference the same memory.
The spatial region represents ownership of a part of memory which ensures that a linear capability is not aliased.
Even though spatial regions give the right to reference part of memory, we still need the world to specify the remainder of the memory that may be referenced by linear capabilities.
To this end, we have shadow regions.
A shadow region is a shadow copy of a spatial region in the sense that it specifies part of memory, but it does not give the right to reference that part of memory.
This does not mean that the memory is necessarily not referenced.
A compatible world may claim a shadow region as spatial which allows the other world to reference the memory (we return to the notion of compatible worlds w.r.t. ownership in Section~\ref{subsubsec:joining-worlds}).
Specifically, we add tags $\spatialo$ and $\spatial$ to the spatial regions:
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \cup \\ 
    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg})  \\
  \end{array} \right.
\]
For readability, we also add a tag $\pure$ to the shared regions:
\[
  \Regionh = \{\pure \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) 
\]

% What parts of memory are linear
We will extend the regions further in Sections~\ref{subsubsec:seals}~and~\ref{subsubsec:ft-and-revocation}, but for now we continue the definitions of the three sub-worlds.
The sub-world $\Worldh$ specifies the heap memory which can be referenced by both linear and normal capabilities, so it should contain both shared and spatial regions.
For this reason, it is defined as
\[
  \Worldh = \RegName \parfun (\Regionh \cup \Regions)
\]
\srccm{} internalizes the \stktokens{} stack, so it can only be referenced by linear capabilities which means that the two stack regions should only have spatial regions.
For instance the $\Worldfs$ is defined as
\[
  \Worldfs = \RegName \parfun \Regions
\]
$\Worlds$ can also only be referenced by linear capabilities, so it should also only have spatial regions.
$\Worlds$ not only models the memory contents of the local stack frames, it also models the call stack that consists of the stack frames.
Conceptually, this means that each of the stack frames is connected with a return point in some code.
In a traditional C calling convention, the code return point would even be stored in the stack frame.
However, with \stktokens{} a stack frame does not contain any information about the corresponding code return point.
Instead, the stack frame and code return point are connected by the capabilities that reference them as they are sealed with the same seal and together then constitute the sealed return pair.
In order to ensure that each call actually returns to the correct point of the code, we must still include the address of the return point in our model.
To this end, each shared region in $\Worlds$ must be paired with a return address:
\[
\Worlds = \RegName \parfun (\Regions \times \Addr)
\]

% Ensuring linearity
The spatial regions add the necessary bookkeeping to the worlds to model linear capabilities.
The logical relation presented in Section~\ref{subsec:logical-relation} uses this bookkeeping to ensure that linear capabilities \emph{uniquely} reference part of memory.

Given a world, we want to be able to express that a capability that is otherwise valid with respect to the world is not linear or indirectly depends on a linear capability.
This is expressed by stripping the world of all its ownership which corresponds to replacing all $\spatialo$ regions with $\spatial$ regions and then requiring that the capability is valid w.r.t.\ this new world.
We refer to this as the shared part of the world and define the function $\purePart{}$ which turns all spatial regions into shadow copies.
\begin{definition}[The shared part of a world]
  \label{def:purePart}
  For any world $W$, we define
  \begin{align*}
    \purePart{W} &\defeq (\purePart{\pwheap},\purePart{\pwpriv},\purePart{\pwfree})\\
    \purePart{W_\var{heap}} &\defeq \lambda r\ldotp
                       \begin{cases}
                         (\spatial,\var{H}) & \text{if } W_\var{heap}(r) = (\spatialo,\var{H})\\
                         W_{\var{heap}}(r) & \text{otherwise}
                       \end{cases}\\
    \purePart{W_\var{call\_stk}} &\defeq \lambda r\ldotp
                       \begin{cases}
                         ((\spatial,\var{H}),\opc) & \text{if } W_\var{call\_stk}(r) = ((\spatialo,\var{H}),\opc)\\
                         W_\var{call\_stk}(r) & \text{otherwise}
                       \end{cases}\\
    \purePart{W_\var{free}} &\defeq \lambda r\ldotp
                       \begin{cases}
                         (\spatial,\var{Hs}) & \text{if } W_{\var{free}}(r) = (\spatialo,\var{Hs})\\
                         W_{\var{free}}(r) & \text{otherwise}
                       \end{cases}
  \end{align*}
\end{definition}
\subsubsection{Seals}
\label{subsubsec:seals}
So far, the world represents a collection of assumptions on the memory contents that value correctness may depend on.
However, value correctness may depend on other assumptions.
Specifically, \stktokens{} has certain assumption on the seals used for return capabilities and closures.
For instance, a return seal must only be used to seal the return pointer of one specific return point.
Therefore, in addition to a relation on memory segments, some regions also carry a \emph{seal interpretation function} that relates the sealables that may be sealed with a given seal.
  \[
    \Seal \parfun \World \fun \Rel{\SealableCaps \times \SealableCaps}
  \]
%% Shared region has a seal indexed invariant that specifies what sealables can be sealed with a specific seal
  In \stktokens{}, once a seal has been used for a specific purpose (e.g.\ for sealing return capability pairs for a specific call site), it can never be reused for a different purpose.
  This is because there may still be copies of return capabilities out there, signed with the seal.
  This situation is similar to the situation for non-linear memory capabilities, so we only allow shared regions to carry seal interpretation functions, as we will see that those regions can never be revoked in future worlds.
\begin{multline*}
  \Regionh = 
  \{\pure \} \times (\World \fun \Rel{\MemSeg\times\MemSeg}) \times \\
  (\Seal \parfun \World \fun \Rel{\SealableCaps \times \SealableCaps})
\end{multline*}
We also refer to the seal interpretation function as the seal invariant, and we will refer to the memory relation as the memory invariant or just invariant when it is unambiguous.

@@

Safety of capabilities should be defined, so that it is monotone with respect to the future world relation.
This means that the system assumptions that capabilities rely on for safety are defined such that capabilities remain safe during execution.
The same holds for the memory invariants stored in the world: capabilities stored in memory should remain valid under allowed changes in system assumptions.
This means we require the world-indexed memory invariants to be monotone in the world.
In other words, a pair of memory segments that are related now will stay related in future worlds.
This changes the spatial regions as follows:
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \cup \\
    \{\spatialo \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \cup \\
    \{\revoked \}
  \end{array} \right.
\]
We make a similar change to $\Regionh$. The seal invariants are monotone as well.

@@

The stack relation also partitions the authority of the world, so it can relate the stack frames in a way that respects linearity.

@@

So far, we have talked about the logical relation as though we define one.
However, we actually define two logical approximations that only differ in the observation relation.
We define a \srccm{} configuration to logically approximate a \trgcm{} configuration when the halting termination of the \srccm{} configuration implies the halting termination of the \trgcm{} configuration.
This also means that \srccm{} configurations that terminate by failing termination are related to any \trgcm{} configuration.
Intuitively, this is because the $\failed$ configuration signals that there was an attempt to break the guarantees of the capability machine.
For instance, a piece of code could have attempted to read from a part of memory it does not have access to, or a callee could have attempted to return out of order.
In both cases, we haven't defined a way to recover from such attempts to break the guarantees, so we are content with failure.
\begin{multline*}
  \lro[\preceq,(\ta,\stkb,\_,\_)] \defeq \\\left\{ \npair{\left(\array{l}\src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})},\\(\ms_T,\reg_T)\endarray\right)} \middle|
    \begin{array}{l}
      \forall i \leq n \ldotp \\
      \quad \src{(\ms_S,\reg_S,\stk_S,\ms_{\stk,S})} \sterm[i]{\ta,\stkb} \\\qquad\Rightarrow (\ms_T,\reg_T) \term\\
    \end{array}
\right\}
\end{multline*}
The step-indexing plays a role here because we are only interested in \srccm{} configurations that terminate in $n$ or fewer steps.
However, if the \srccm{} configuration terminates successfully in $n$ steps, then the \trgcm{} configuration should just terminate in any number of step (possibly more than $n$ steps).
For the most part, it would make sense to require the \trgcm{} configuration to terminate in the same amount of steps as the \srccm{} configuration as they run in lockstep for most of the computation.
However, when it comes to calls and returns, the two configurations stop running in lockstep.
The \srccm{} configuration handles calls and returns in one step whereas \trgcm{} configurations need to execute each instruction of the call preparation as well as the return code.

@@

In Figure~\ref{fig:value-relation}, we have sketched the two value relations.
This shows that for the most part, words on \srccm{} are related to words on \trgcm{} that are syntactically identical.
The only exception is stack pointers on \srccm{} that are related to linear capabilities on \trgcm{}.
Note that the return pointers of \srccm{} are not related to anything as it is never safe for any program, trusted or not, to have them.
The \srccm{} return pointers should only occur under a return seal, and they should only be used in a jump in which case the \srccm{} semantics transforms them to the capabilities they correspond to.

@@

In our result, we assume well-formed components which puts certain syntactic constraints on the components.
We also have the semantic assumption that trusted components are reasonable.
Both assumptions need to be captured in the logical relation in order for us to rely on them.
To this end, we define a \emph{code region} which captures the syntactic and semantic assumptions we make on components.
The memory invariant of the code region is defined as
\begin{multline*}
  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} =\\
  \left\{\npair{\arraycolsep=0pt\left(\array{l}\code \uplus \mspad,\\ \code \uplus \mspad\endarray\right)} \middle|
    \begin{array}{l}
    \dom(\code) = [\baddr,\eaddr] \wedge ([\baddr - 1, \eaddr + 1] \subseteq \ta
      \wedge\\
      \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss \wedge \trust = \trusted) \vee \\
      \quad ([\baddr-1,\eaddr+1]\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
      \mspad = [\baddr-1 \mapsto 0] \uplus [\eaddr + 1 \mapsto 0]\wedge\\
      \sigrets,\sigcloss,\ta \vdash_{\mathrm{comp-code}} \code \wedge\\
      \forall a \in \dom(\code)\ldotp\\
      \quad\npair{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
    \end{array}
  \right\}
\end{multline*}
The code region is more restrictive than the standard region.
It only allows one memory segment, namely $\code$ padded with zeroes that make sure that two capabilities cannot be spliced to cause unintended control-flow.
We use the relation to reason about trusted components (well-formed and reasonable) as well as untrusted components (well-formed).
The assumptions we can make on the code depend on whether it is part of a trusted or untrusted component.
This is captured by requiring the contents of the code memory to be in the trusted or untrusted value relation depending on the trustworthiness of the code.
That is, if all the code memory addresses are in the trusted address space and the seals are from the global seals, then the component is trusted.
On the other hand, if the code memory addresses are disjoint from the trusted addresses and there are no return seals, then the component is untrusted.
In either case, the words should be in the value relation with respect to the $\purePart{}$ of the world which means that the code memory cannot contain linear capabilities.

@@

% Closure seals
Unlike return seals, both trusted and untrusted components can have closure seals.
For untrusted components (components with their code address space disjoint from the trusted address space), we allow everything in the untrusted value relation to be sealed.
Intuitively, untrusted components are assumed to have access to words from the untrusted value relation, and we cannot know how the words are used, so we need to assume that an untrusted component may seal untrusted words.
Trusted components only use closure seals for sealed capability pairs that represent actual closures.
The code capability for a closure must point to the code memory because it is the only part of memory that is executable.
Untrusted components cannot safely have a capability for a trusted component's code (it could be used to read return capabilities or start execution in the middle of a call), so capabilities for the code memory of a trusted component is in the trusted value relation.
While it is not safe to give a bare capability for a trusted components code memory, it can be perfectly safe to give a sealed capability for a trusted components code.
For this reason, the seal invariant allows executable capabilities from the trusted value relation to be sealed with a closure seal.

@@

The world can be seen as an authority specification which means that it dictates what kind of capabilities can address a certain part of memory.
Specifically, linear capabilities can only address memory governed by a $\spatialo$ region, and normal capabilities can only address memory governed by a $\pure$ region.
All the permission based conditions we define project the regions that the capability may address from the world.
The addressable $\addressable{}$ function takes care of the projection:
\[
  \addressable{\lin,W} \defeq
  \begin{cases}
    \{ r \mid W(r) = (\pure,\_) \} & \text{if $\lin = \normal$} \\
    \{ r \mid W(r) = (\spatialo,\_) \}  & \text{otherwise (i.e. $\lin = \linear$)} \\
  \end{cases}
\]

We capture the essence of what it means for a capability with read permission to be safe in the condition $\readCond[]{}$.
% each of the regions upper bounded by std region
The main purpose of $\readCond[]{}$ is to make sure that only safe words can be read from the memory governed by a read capability.
This is done by putting an upper bound on what requirements an invariant can impose on the memory segments governed by the capability.
In particular a region that governs the memory a read capability has access to can at most allow safe values to be read.
Without this requirement, a read capability could potentially be used to break memory invariants if it were used to read capabilities that has the authority to break memory invariants.
The read condition is defined as follows
\[
  \readCond{\lin,W} = \left\{ \npair{A} \middle| 
    \begin{array}{l}
      \exists S \subseteq \addressable{\lin, \pwheap} \ldotp \\
      \quad \exists R : S \fun \powerset{\nats} \ldotp\\
      \qquad \biguplus_{r\in S} R(r) \supseteq A \wedge\\
      \qquad (\lin = \linear \Rightarrow \forall r \in S \ldotp |R(r)|  = 1) \wedge\\
      \qquad \forall r \in S \ldotp \pwheap(r).H \nsubeq \stdreg{R(r),\gc}{\pur}.H
    \end{array}
  \right\}
\]
% splicing and what part of the range each region governs
The $\readCond[]{}$ is compatible with all the operations that can be performed on capabilities.
This means that if two capabilities for which $\readCond[]{}$ holds are spliced together, we can establish that $\readCond[]{}$ holds for the resulting capability.
To support this, we require the presence of a set of regions $S$ that governs the addresses the capability has authority over rather than just a single region.
If we need to establish the $\readCond[]{}$ after a splice, we can simply use the union of the regions that witnessed the $\readCond[]{}$ of the two individual capabilities.
We also need to support splitting which is no problem for normal capabilities as the same $\pure$ region can be used to establish the $\readCond[]{}$ for multiple normal capabilities.
On the other hand, a $\spatialo$ region can only be used to establish the $\readCond[]{}$ for one linear capability because the ownership of a spatial region can only go to one world when splitting the ownership.
Nonetheless, we need to support arbitrary splitting of linear capabilities, which means that $\readCond[]{}$ must make sure that the necessary regions are in the world to argue that the result of a split preserves $\readCond[]{}$.
This is why $\readCond[]{}$ requires all regions to only govern one address when the capability is linear.
This means that after a split, the authority of the regions for the bottom half of the split can go to one capability and the remaining regions can go to the top half.

@@

Similarly to $\readCond[]{}$, we define a condition that captures the essence of what it means for a capability with writer permission to be safe.
We call this condition $\writeCond[]{}$.
A capability with write permission can be used to write to memory.
The question is what can we safely allow to be written to memory without any memory invariants being broken.
The answer to this is anything - even words that are unsafe.
Say, you manage to write something that can break memory invariants, then it would not be possible to read it back again as write permission, generally speaking, does not entail read permission.
If the capability had read permission, then $\readCond[]{}$ would make sure that the word would have to be safe\footnote{It should not be possible to obtain a capability that can be used to break invariants. After all, if such a capability was obtained, memory invariants could be broken.
  However, the $\writeCond[]{}$ tries to capture the essence of safety and in principle it is safe to write an unsafe capability that cannot be read back.}.
It should always be possible to write safe values, so we impose this as a lower bound.
